Large-scale feature selection with Gaussian mixture models for the classification of high dimensional remote sensing images

### 2016-12-15 response.tex

\begin{revbox}
  Following the question 3, why choosing the optimal number of variables based on the increase of criterion instead of the highest value?
  \begin{resbox}
    We also considered this possibility and it is a choice available in the code of the OTB module we provide. But taking the maximum appears not to be the optimal choice in the sense that the algorithm selects sometimes more variables than necessary. Sometimes, at a given point, the best new selected variable is one which contains no information (for example one whose contribution is thresholded by EPS\_FLT) because all others decrease the criterion value. If it happens several times, it can lead to a final drop of the criterion. The algorithm then tries to counterbalance the drop and regains the original maximum but can be slightly higher. Figure~\ref{fig:crit-evol} illustrated this case.
    The problem is that we introduce a few irrelevant features and we want to avoid that if possible. This case appears more often with the second dataset but we choose to use the same method for both dataset.
  \end{resbox}
\end{revbox}

\begin{figure}[!t]
  \centering
  \begin{tikzpicture}
    \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
      \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
    \end{axis};
  \end{tikzpicture}
  \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Potsdam dataset with 1000 samples by class.\label{fig:crit-evol}}
\end{figure}

\begin{revbox}
  In the experiment with Aisa dataset, although several numbers of training samples are tested, the number of training samples seems carefully selected by the authors to show the cases where the GMM classifiers perform better. I guess when a larger number of samples are trained, the random forest will have a higher classification rate. And when a smaller number of samples are used, the KNN will have a shorter training time. The GMM only wins in the classification time regardless of the sample number. It’ll be good to show these cases.
  \begin{resbox}
    It is true that we can try with fewer samples. We tried with 50 samples by class and it leads to the same behavior than with 250 samples (see Table~\ref{tab:aisa-otbsimu}).

    But in the case of a larger training set is not possible to test with the chosen validation method. We use an equal number of samples in each class and 50\% of the samples are used for test. Thus, as the less represented class contains 2107 samples, we are limited to 1053 samples by class with this dataset. Nevertheless, the second dataset illustrates the case and we also add the following comment in the paper to make it clear that RF could perform better with more samples:
    \emph{Random Forest has similar performance in term of classification rate with 1000 samples \rev{ and one could expect to get a better classification rate with RF if more samples were available}.}
  \end{resbox}
\end{revbox}

\begin{table}[!t]
    \centering
    \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
    \begin{tabular}{lcccc}\toprule
         & \multicolumn{4}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-5}
        \# samples by class & 50 & 250 & 500 & 1000 \\ \midrule

        GMM SFS kappa &  0.612 (0.028) & {\bfseries 0.678 (0.029)} & {\bfseries 0.687 (0.029)} & {\bfseries 0.699 (0.028)} \\
        GMM SFS JM &     0.622 (0.026) & {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
        GMM SFFS JM &    0.622 (0.026) & {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
        GMM ridge &      0.597 (0.035) & 0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
        KNN &            0.481 (0.054) & 0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
        Random Forest &  0.555 (0.031) & 0.645 (0.026) & 0.673 (0.023) & {\bfseries 0.693 (0.023)} \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{revbox}
  Correct some typos, for example:
  \begin{enumerate}
    \item In Proposition 2, in the sentence after the covariance matrix equation, I guess it should be ``u\_c'' instead of ``v\_c''.
    \item Grammar mistakes such as ``a buildings''.
  \end{enumerate}
  \begin{resbox}
    % Thank you for the careful reading.
    NB:

    no $v_c$ after prop 2...

    building corrected. Need to correct english at the end
  \end{resbox}
\end{revbox}

### 2016-12-16 response.tex

\begin{revbox}
In Algorithm 3, a threshold EPS\_FLT on eigenvalues is used to prevent numerical instability. It’s not clear how the authors select the proper value for this threshold. And how does this threshold influence the classification accuracy?

  \begin{resbox}
    In order to influence as less as possible the classification accuracy, the threshold has been set as small as possible. The smallest possible float number is define by the computer architecture (in our case 64bits) and C++ specifications. The C++ documentation states that epsilon is the \emph{difference between 1 and the least value greater than 1 that is representable}(\url{http://www.cplusplus.com/reference/cfloat/}).
    This thresholding can be seen as a ridge regularization when applied to eigenvalues. 
  \end{resbox}

\end{revbox}

\begin{revbox}
  The authors show the experimental results on two different types of hyper-dimensional images, which is very good. But I have some doubts on the construction of the second high dimensional cube where 191 features are constructed based on the low dimensional RGBIR images. Are these features simply used for the purpose of increasing the data dimensionality? Or are there benefits in terms of classification accuracy?
  \begin{resbox}
    The interest of this second data cube is first to test the algorithm in a different context. In particular, the algorithm applied to this second dataset should be seen as a method to learn the most relevant kind of features and the best parameters to extract those features.

    In the first dataset, only spectral features has been considered and in the particular context of hyperspectral imaging which imply contiguous features highly correlated.

    (reponse bof à revoir et developper mais c'est un debut)
  \end{resbox}
\end{revbox}

@@

\begin{revbox}
  Correct some typos, for example:
  \begin{enumerate}
    \item In Proposition 2, in the sentence after the covariance matrix equation, I guess it should be ``u\_c'' instead of ``v\_c''.
    \item Grammar mistakes such as ``a buildings''.
  \end{enumerate}
  \begin{resbox}
    % Thank you for the careful reading.
    NB:

    no $v_c$ after prop 2...

    building corrected. Need to correct english at the end
  \end{resbox}
\end{revbox}

\section{Reviewer 2}
\begin{revbox}
  In section III-A, 5 criterion functions are proposed. With 2 feature selection methods (SFS and SFFS), there should be 10 possible configurations. However, in the experimental parts (page 14, line 21-24), results of only three configurations are shown. The authors said “other configurations have been investigated and performs either equally or lower in terms of classification accuracy.” I would suggest that the authors either show the results of more configurations or not introducing those unrelated criterion functions.
  \begin{resbox}
    see Table~\ref{tab:aisa-otbsimu-all}
  \end{resbox}
\end{revbox}

    \begin{table}[!t]
        \centering
        \caption{Average classification accuracy 20 trials with all configurations of criterion and method (standard deviation in parenthesis).\label{tab:aisa-otbsimu-all}}
        \begin{tabular}{lccc}\toprule
             & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
            \# samples by class & 250 & 500 & 1000 \\ \midrule

            GMM SFS JM &    0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
            GMM SFS divKL &    &&\\
            GMM SFS kappa & 0.678 (0.029) & 0.687 (0.029) & 0.699 (0.028) \\
            GMM SFS OA &    &&\\
            GMM SFS F1mean &    &&\\
            GMM SFFS JM &   0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
            GMM SFFS divKL &   &&\\
            GMM SFFS kappa &   &&\\
            GMM SFFS OA &   &&\\
            GMM SFFS F1mean &   &&\\
            \bottomrule
        \end{tabular}
    \end{table}

\begin{revbox}
  Only Cohen’s Kappa is utilized to analyze the classification accuracy. The experimental results would be more persuasive if other criterion functions are utilized as well. I am interested to see how different methods’ performance varies under different criterion functions.
  \begin{resbox}
    see Table~\ref{tab:aisa-otbsimu-othereval}
  \end{resbox}
\end{revbox}

    \begin{table}[!t]
        \centering
        \caption{Average classification accuracy 20 trials evaluated with overall accuracy and mean of F1-score (standard deviation in parenthesis).\label{tab:aisa-otbsimu-othereval}}
        \begin{tabular}{lcccccc}\toprule
             & \multicolumn{3}{c}{\bfseries Overall Accuracy} & \multicolumn{3}{c}{\bfseries F1-mean} \\ \cmidrule{2-7}
            \# samples by class & 250 & 500 & 1000  & 250 & 500 & 1000 \\ \midrule

            GMM SFS kappa & & & & & & \\
            GMM SFS JM &    & & & & & \\
            GMM SFFS JM &   & & & & & \\
            GMM ridge &     & & & & & \\
            KNN &           & & & & & \\
            Random Forest & & & & & & \\
            \bottomrule
        \end{tabular}
    \end{table}

@@

\begin{revbox}
  A reference or an explanation should be made in Page 1, line 36 regarding the curse of dimensionality
  \begin{resbox}

  \end{resbox}
\end{revbox}

\section{Reviewer 3}

\begin{revbox}
  Although the method seems promising, there are some significant issues the authors should address to. The key novelty and contribution of this technique is unfortunately unclear. The authors dedicate a lot of space in explaining already known mathematical formulations of the Gaussian Mixture Models and the feature selection techniques, in comparison with their proposed scheme.
  \begin{resbox}
    We totally agree that some mathematical formulations we used are already known. Nevertheless, we think that these formulations in this particular context has not been exploited so far. The results of our experiment shows that the proposed algorithm has the advantage to be very efficient in term of processing time with a fairly reasonable cost in term of classification accuracy. It has to be noticed that compared to similar methods using Gaussian Mixture model, no additional hypothesis are made on the covariance matrix which is tractable only because of the efficient implementation.
  \end{resbox}
\end{revbox}

@@

\begin{revbox}
  Additionally, critical references are missing, such as:
  \begin{enumerate}
  \item K. Karalas, G. Tsagkatakis, M. Zervakis, and P. Tsakalides, "Land Classification Using Remotely Sensed Data: Going Multi-Label," IEEE Transactions on Geoscience and Remote Sensing, 2016, doi: 10.1109/TGRS.2016.2520203
  \item K. Karalas, G. Tsagkatakis, M. Zervakis, and P. Tsakalides. "Deep learning for multi-label land cover classification", in SPIE Remote Sensing, Toulouse
  \end{enumerate}

  \begin{resbox}
    After reading with interest this two references, we fail to see the link between these publications and our submission except for the type of data used to test algorithms.

    First reference reviews ways to assign several labels to one sample and so performs spectral unmixing. Second reference focus on the same topic but develop a deep learning approach. Although very interesting, we do not address the subject of multi-labeling in our work.

    According to these observations, we choose to respectfully disagree with the reviewer and take the decision not to add a digression to cite these references.
  \end{resbox}
\end{revbox}

### 2017-01-10 paper.tex
add data file to plot derivate of criterion
update

% \documentclass[journal,peerreview,onecolumn]{IEEEtran}
\documentclass[journal]{IEEEtran}
\usepackage{etex}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{./Fig/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{url}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{empheq}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{minted}\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1.2cm{\vfil
      \hbox to 1.2cm{\hfil\parbox{1.4cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

@@

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}

@@

In the spectral or temporal domain, a pixel is represented by a vector
for   which  each   component  corresponds   to  a   spectral/temporal
measurement.   The size  of  the  vector is  therefore  the number  of
spectral  or temporal  measurements.  For  hyperspectral images,  this
number is  typically about several  hundreds while for  the Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}. This  is a  major drawback  in many  remote sensing
application since it  is difficult to collect a large  and an accurate
ground-truth.   An intensive  work has  been performed  in the  remote
sensing community  to build accurate classifiers  for high dimensional
images.  Bayesian models~\cite{book:landgrebe}, feature extraction and
feature   reduction   techniques~\cite{book:landgrebe,DR:guided:tour},
random   forest~\cite{1396322},  neural   networks~\cite{5411821}  and
kernel methods~\cite{kernel:methods:rs} have been investigated for the
classification of such images.

The volume of the data is  increasing dramatically with respect to the
number of measurement  per pixel. The data volume  of an hyperspectral
image  is  typically several  hundreds  of  Gigabytes per  acquisition
($\approx$  300km$^2$). Multitemporal  data are  now available  freely
from  internet  stream  (see  for instance  the  Copernicus  data  hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of parallel approach (multi-core, GPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  National
Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  conveniently proposed  to the  users/developpers through  several
modules ``\emph{ready to use}''.

@@

        \begin{prop}[Mean update for cross-validation]
            \label{eq:update-cv1}
            \begin{equation*}
                \boldsymbol{\hat{\mu}}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\hat{\mu}}_c^{n_c} - \nu_c \boldsymbol{\hat{\mu}}_c^{\nu_c}}{n_c - \nu_c} \nonumber
            \end{equation*}
        \end{prop}
        \begin{prop}[Covariance matrix update cross-validation]
            \label{eq:update-cv2}
            \begin{align*}
              \boldsymbol{\hat{\Sigma}}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl\{ (n_c-1) \boldsymbol{\hat{\Sigma}}_c^{n_c} - (\nu_c-1)\boldsymbol{\hat{\Sigma}}_c^{\nu_c} \nonumber \\
                                                        &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})(\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})^t \biggr\} \nonumber
            \end{align*}
        \end{prop}
        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exponents on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.

@@

    where $\boldsymbol{\Lambda}^{(k)}_c$  and $\mathbf{P}^{(k)}_c$ are
    the diagonal  matrix of eigenvalues  of the covariance  matrix and
    the    orthonormal   matrix    of   corresponding    eigenvectors,
    respectively.  To prevent  numerical instability,  the eigenvalues
    are thresholded to a given  value EPS\_FLT. In our implementation,
    EPS\_FLT is set to the floating machine precision.

@@

    In order to increase the dimensionality of the data, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item Fifteen Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
        \item Morphological profile build on each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
        \item Attribute profile build on each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Attribute profile build on each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Textural features for  each channel with neighborhood of
          19x19 pixels:  mean, standard  deviation, range  and entropy
          (16 features)~\cite{otb}.
    \end{itemize}
    The normalized DSM and the raw  RGBIR image are added to these 191
    features and  then stacked to  create a  new image  with 196
    bands. %  The  training is  made with  tile 5\_11  and test  with two
    % tiles: 5\_12 which very similar with  tile 5\_11 and 3\_10 which a
    % further area a bit less urban and more industrial.

@@

    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
    \begin{itemize}
    \item One with forward selection and JM distance as criterion (GMM SFS JM);
    \item One with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
    \item One with floating forward selection and JM distance as criterion (GMM SFFS JM).
    \end{itemize}
    Other configurations  have been  investigated and  performs either
    equally     or     lower     in    terms     of     classification
    accuracy~\cite{al:report}. For the sake of clarity, only the three
    aforementioned configurations are discussed here.

@@

    For SFS and  SFFS selection, the number of variables  to select is
    set to  30 for the  Aisa dataset and  60 for the  Potsdam dataset.
    After  the selection  procedure, the  optimal number  of extracted
    variables is selected as follow.  Rather than selecting the number
    of variables corresponding to the  highest value of the criterion,
    the number of  retained variables is set when  the criterion stops
    to increase significantly.  It is  found by computing the discrete
    derivative of the criteria.  See Figure~\ref{fig:crit-evol} for an
    example.

@@

    The    classification   rate    is    presented   using    Cohen's
    kappa. Processing  time has been  evaluated on a  desktop computer
    with  8Gb of  RAM  and  Intel(R) Core(TM)  i5-3570  CPU @  3.40GHz
    $\times$ 4 processors.

@@

    \begin{table}[!t]
        \centering
        \caption{Mean processing time for training and classification for results in Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
        \begin{tabularx}{\columnwidth}{l*{6}{>{\centering\arraybackslash}X}}
            \toprule
             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} \\ \cmidrule{2-7}
            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule

            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 5.2} & {\bfseries 5.2} & {\bfseries 5.5} \\
            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 5.7} & {\bfseries 5.7} & {\bfseries 5.9} \\
            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 5.0} & {\bfseries 5.0} & {\bfseries 5.4} \\
            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530 \\
            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887 \\
            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 \\
            \bottomrule
        \end{tabularx}
    \end{table}

@@

    The difference with the second best classifier, Random Forest, appears to be significant when using 250 and 500 samples. Random Forest has similar performance in term of classification rate with 1000 samples and one could expect to get a better classification rate with RF if more samples were available. The GMM classifier with ridge regularization and the KNN classifier are both outperformed.

>>>>

% \documentclass[journal,peerreview,onecolumn]{IEEEtran}
\documentclass[journal,10pt]{IEEEtran}
\usepackage{etex}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{./Fig/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{url}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{empheq}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{minted}\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1.2cm{\vfil
      \hbox to 1.2cm{\hfil\parbox{1.4cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

@@

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}
\newcommand{\rev}[1]{\textcolor{magenta}{\emph{#1}}}

@@

In the spectral or temporal domain, a pixel is represented by a vector
for   which  each   component  corresponds   to  a   spectral/temporal
measurement.   The size  of  the  vector is  therefore  the number  of
spectral  or temporal  measurements.  For  hyperspectral images,  this
number is  typically about several  hundreds while for  the Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}~\cite{hughes}. This  is a  major drawback  in many  remote sensing
application since it  is difficult to collect a large  and an accurate
ground-truth.   An intensive  work has  been performed  in the  remote
sensing community  to build accurate classifiers  for high dimensional
images.  Bayesian models~\cite{book:landgrebe}, feature extraction and
feature   reduction   techniques~\cite{book:landgrebe,DR:guided:tour},
random   forest~\cite{1396322},  neural   networks~\cite{5411821}  and
kernel methods~\cite{kernel:methods:rs} have been investigated for the
classification of such images.

The volume of the data is  increasing dramatically with respect to the
number of measurement  per pixel. The data volume  of an hyperspectral
image  is  typically several  hundreds  of  Gigabytes per  acquisition
($\approx$  300km$^2$). Multitemporal  data are  now available  freely
from  internet  stream  (see  for instance  the  Copernicus  data  hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of parallel approach (multi-core, GPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  Space
Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  conveniently proposed  to the  users/developpers through  several
modules ``\emph{ready to use}''.

@@

        \begin{prop}[Mean update for cross-validation]
            \label{eq:update-cv1}
            \begin{equation*}
                \boldsymbol{\hat{\mu}}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\hat{\mu}}_c^{n_c} - \nu_c \boldsymbol{\hat{\mu}}_c^{\nu_c}}{n_c - \nu_c} \nonumber
            \end{equation*}
        \end{prop}
        \begin{prop}[Covariance matrix update cross-validation]
            \label{eq:update-cv2}
            \begin{align*}
              \boldsymbol{\hat{\Sigma}}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl\{ (n_c-1) \boldsymbol{\hat{\Sigma}}_c^{n_c} - (\nu_c-1)\boldsymbol{\hat{\Sigma}}_c^{\nu_c} \nonumber \\
                                                        &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})(\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})^t \biggr\} \nonumber
            \end{align*}
        \end{prop}
        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exponents on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\mu}_c$ denotes the set of samples used to compute them.

@@

    where $\boldsymbol{\Lambda}^{(k)}_c$  and $\mathbf{P}^{(k)}_c$ are
    the diagonal  matrix of eigenvalues  of the covariance  matrix and
    the    orthonormal   matrix    of   corresponding    eigenvectors,
    respectively.  To prevent numerical instability, \rev{non strictly
      positive eigenvalues are thresholded to a fixed value EPS\_FLT}.
    In our  implementation, EPS\_FLT  is set  to the  floating machine
    precision.

@@

    In order to increase the dimensionality of the data,

    \rev{Conventionally,  the following  features are  extracted using  the
    RGBIR  images in  order to  increase the  classification accuracy,
    similarly to \cite{tuia2015multiclass}:}
    \begin{itemize}
        \item Fifteen Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
        \item Morphological profile build on each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
        \item Attribute profile build on each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Attribute profile build on each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Textural features for  each channel with neighborhood of
          19x19 pixels:  mean, standard  deviation, range  and entropy
          (16 features)~\cite{otb}.
    \end{itemize}
    The normalized DSM and the raw  RGBIR image are added to these 191
    features and  then stacked to create  a new image with  196 bands.
    \rev{The  resulting  data cube  is  therefore  high-dimensional.}

@@

    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
    \begin{itemize}
    \item One with forward selection and JM distance as criterion (GMM SFS JM);
    \item One with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
    \item One with floating forward selection and JM distance as criterion (GMM SFFS JM).
    \end{itemize}
    Other configurations  have been  investigated and  performs either
    equally     or     lower     in    terms     of     classification
    accuracy~\cite{al:report}. For the sake of clarity, only the three
    aforementioned configurations are discussed here \rev{and the results for all other configurations is available in the supplementary material}.

@@

    For SFS and  SFFS selection, the number of variables  to select is
    set to  30 for the  Aisa dataset and  60 for the  Potsdam dataset.
    After  the selection  procedure, the  optimal number  of extracted
    variables is selected as follow.  Rather than selecting the number
    of variables corresponding to the  highest value of the criterion,
    the number of  retained variables is set when  the criterion stops
    to  increase significantly.   \rev{It  is found  by computing  the
      discrete derivative  of the criteria between  two iterations and
      normalizing it  by its  maximum value.   The number  of selected
      features  corresponds to  the  last iteration  before the  value
      drops    below    $10^{-3}$     for    all    datasets}.     See
    Figure~\ref{fig:crit-evol} for an example.

@@

    The    classification   rate    is    presented   using    Cohen's
    kappa \rev{but scores computed with overall accuracy and mean of f1-score are available in supplementary material}. Processing  time has been  evaluated on a  desktop computer
    with  8Gb of  RAM  and  Intel(R) Core(TM)  i5-3570  CPU @  3.40GHz
    $\times$ 4 processors.

@@

    \begin{table*}[!t]
        \centering
        \caption{Mean processing time for training and classification for results in Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
        \begin{tabularx}{0.8\textwidth}{l*{9}{>{\centering\arraybackslash}X}}
            \toprule
             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} & \multicolumn{3}{c}{\bfseries \# of selected features} \\ \cmidrule{2-10}
            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule

            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 5.2} & {\bfseries 5.2} & {\bfseries 5.5} & 11.95 & 12    & 12.05\\
            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 5.7} & {\bfseries 5.7} & {\bfseries 5.9} & 11.95 & 12    & 12.05\\
            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 5.0} & {\bfseries 5.0} & {\bfseries 5.4} & 21.45 & 24.35 & 27.05\\
            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530    & & & \\
            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887    & & & \\
            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 & & & \\
            \bottomrule
        \end{tabularx}
    \end{table*}

@@

    The difference with the second best classifier, Random Forest, appears to be significant when using 250 and 500 samples. Random Forest has similar performance in term of classification rate with 1000 samples \rev{and one could expect to get a better classification rate with RF if more samples were available}. The GMM classifier with ridge regularization and the KNN classifier are both outperformed.
